
\section{Tautness}
\subsection{Tautness for vf-categories}

Let $\X$ be a vf-category. Note first that if a vf-arrow $u \colon p \vuto\tau q_{\phi t}$ is the thickening of $v \colon p \vuto\sigma q_s$ with respect to some cover $\phi \colon \tau \epi \sigma$, i.e.\ $u = \phi^* v$, then in particular $\psi_1^* u= \psi_2^* v$ for any pair of $\FF$-arrows $\psi_1 ,\psi_2$ such that $\phi \circ \psi_1 = \phi\circ\psi_2$.  

\begin{defn}
    For a cover $\phi \colon \tau \epi \sigma$, a vf-arrow $u \colon p \vuto{\tau} q_{\phi t}$ in $\X$ is \emph{$\phi$-reducible} (or \emph{reducible along $\phi$}) if $\psi_1^* u = \psi_2^*u$ for every pair of parallel $\FF$-arrows $\psi_1,\psi_2$ such that $\phi \circ \psi_1 = \phi\circ\psi_2$. 
\end{defn}
\begin{rmk}
    Equivalently, to see that $u \colon p \vuto{\tau} q_{\phi t}$ is $\phi$-reducible, it suffices to see that $\pi_1^* u = \pi_2^* u$ where $\pi_1,\pi_2 \colon \tau \times_\sigma \tau \rightrightarrows \tau$ is the kernel pair of $\phi$.
\end{rmk}

As we now define, in a \emph{taut} vf-category, reducible vf-arrows can be {uniquely reduced}.

\begin{defn}\label{def:taut}
    A vf-category $\X$ is \emph{taut} if for every $\phi$-reducible vf-arrow $u \colon p \vuto\tau q_{\phi t}$ there exists a unique vf-arrow $u_0 \colon p \vuto\sigma q_s$ such that $u = \phi^* u_0$.
    % \begin{enumerate}
    %     \item for every $\phi$-reducible vf-arrow $u \colon p \vuto\tau q_{\phi t}$ there exists a unique vf-arrow $u_0 \colon p \vuto\sigma q_s$ such that $u = \phi^* u_0$;
    %     % \item the canonical map $\X(p, (q_t)_{t:\coprod_{i} \sigma_i} ) \to \prod_{i} \X(p, (q_s)_{s:\sigma_i})$ defined by specialization along each coproduct injection $\sigma_i \hookrightarrow \coprod_i \sigma_i$ is a bijection. 
    % \end{enumerate} 
\end{defn}

\begin{ex}\label{ex:top-sp-are-taut-vf-posets}
    Topological spaces, seen as vf-posets as in \Cref{ex:top-sp-as-vf-posets}, are taut. 
\end{ex}

\begin{lem}\label{lem:set-is-taut}
The vf-category $\Set$ is taut.
\begin{proof}
    Fix a set $X$, an $S$-family of sets $(Y_s)_{s\in S}$, and a filter $\sigma$ on $S$. For a cover $\phi \colon \tau \epi \sigma$, let $u \colon X \to \int_{t:\tau} Y_{\phi t}$ be a function into the reduced product of the family $(Y_{\phi t})_{t\in T}$ which, seen as a vf-arrow, is $\phi$-reducible. We need to show that $u$ factors uniquely through the reindexing function $\phi^* \colon \int_{s:\sigma} Y_s \to \int_{t:\tau} Y_{\phi t}$.
    \[\begin{tikzcd}
        X & {\int_{t:\tau}Y_{\phi t}} \\
        {\int_{s:\sigma}Y_{s}}
        \arrow["u", from=1-1, to=1-2]
        \arrow["{u_0}"', dashed, from=1-1, to=2-1]
        \arrow["{\phi^*}"', from=2-1, to=1-2]
    \end{tikzcd}\]
    Let $\rho \coloneqq \tau\times_\sigma\tau$ be the kernel pair of $\phi$, with its two projections $\pi_1,\pi_2 \colon \rho \rightrightarrows \tau$, and denote by $\gamma \colon \rho \epi \sigma$ the composite $\phi\pi_1=\phi\pi_2$. Since $u$ is $\phi$-reducible, we know that $\pi_1^* u = \pi_2^*u$, so that the claim follows if we show that the following diagram is an equalizer in $\Set$:
\[\begin{tikzcd}
	{\int_{s:\sigma} Y_s} & {\int_{t:\tau}Y_{\phi t}} & {\int_{r:\rho}Y_{\gamma r}}
	\arrow["{\phi^*}", from=1-1, to=1-2]
	\arrow["{\pi_2^*}"', shift right, from=1-2, to=1-3]
	\arrow["{\pi_1^*}", shift left, from=1-2, to=1-3]
\end{tikzcd}\]
To see this, recall first that elements of the reduced product $\int_{s:\sigma}Y_{s}$ are in bijection with $\FF$-arrows $\sigma \to \sum_{s\in S} Y_s$ lying above $S$, as shown below. 
\[\begin{tikzcd}
	&&& {\sum_{s\in S}Y_s} \\
	\rho & \tau & \sigma & S
	\arrow[from=1-4, to=2-4]
	\arrow["{\pi_2}"', shift right, from=2-1, to=2-2]
	\arrow["{\pi_1}", shift left, from=2-1, to=2-2]
	\arrow["\phi", two heads, from=2-2, to=2-3]
	\arrow[dashed, from=2-3, to=1-4]
	\arrow[hook, from=2-3, to=2-4]
\end{tikzcd}\]
The claim follows then recalling that $\phi$ is the coequalizer of its kernel pair since every epimorphism in $\FF$ is effective.
\end{proof}
\end{lem}

\subsection{Vf-fully-faithfulness and subspaces}In this subsection, we introduce some properties of vf-functors which will be central later on, but which are also of independent interest. We will see how they behave as expected under the hypothesis of tautness. Moreover, we will employ them to introduce the notion of \emph{(open) subspaces} of a vf-category.

\begin{defn}
    A vf-functor $F \colon \X \to \mathscr Y$ is \emph{vf-faithful} (resp.\ \emph{vf-fully-faithful}) if, for each point $p$, $S$-family of points $(q_s)_{s\in S}$, and filter $\sigma$ on $S$, the function
    \[ \Hom^{\X}_\sigma (p, q_s) \to \Hom^{\Y}_\sigma (F(p), F(q_s)) \]
    is injective (resp.\ bijective).
\end{defn}

\begin{defn}
    A vf-functor $F \colon \X \to \mathscr Y$ is \emph{vf-full} if for each vf-arrow $u \colon F(x) \vuto{s:\sigma} F(y_s)$ in $\Y$ there exists a cover $\phi \colon \tau \epi \sigma$ in $\FF$ and a vf-arrow $v \colon x \vuto{t:\tau} y_{\phi t}$ in $\X$ such that $F(v) = \phi^*(u)$.
\end{defn}

\gabrielnote{Maybe fist define vf-full, then the factorization vf-full surj-on-ob / ff, then define the topological reflection from the factorization of the morphism into the terminal.}

\begin{lem}\label{lem:vf-fully-faithful-equiv}
    Let $F \colon \X \to \Y$ be a vf-functor between taut vf-categories. Then, the following are equivalent:
    \begin{enumerate}
        \item $F$ is vf-full and vf-faithful;
        \item $F$ is vf-fully-faithful.
    \end{enumerate}
\end{lem}
\begin{proof}
    Clearly, $(2)\Rightarrow (1)$. Conversely, suppose $F$ is vf-full and vf-faithful. To show that it is vf-fully-faithful we need to show that, for each $p \in \X_0$ and each $(q_s)_{s:\sigma} \in \X_0^\sigma$, the function
    \[ \X( p, (q_s)_{s:\sigma}) \to \Y (F(p), (F(q_s))_{s:\sigma}) \]
    is surjective, so let $u \colon F(p) \vuto\sigma F(q_s)$ be a vf-arrow in $\Y$. By vf-fullness of $F$, there exist a cover $\phi \colon \tau \epi \sigma$ and a vf-arrow $v \colon p \vuto\tau q_{\phi t}$ in $\X$ such that $F(v)= \phi^* u$. Now, for any $\psi_1,\psi_2 \colon \lambda\to \tau$ equalizing $\phi$, we have 
    \[F(\psi_1^*v) = \psi_1^*F(v) = \psi_1^*\phi^*u = \psi_2^*\phi^*u = \psi_2^*F(v) = F(\psi_2^*v)\]
    and hence $\psi_1^*v = \psi_2^*v$ by vf-faithfulness of $F$. In other words, $v$ is $\phi$-reducible: since $\X$ is taut, it follows that there is a (unique) vf-arrow $v_0 \colon p \vuto\sigma q_s$ such that $v = \phi^* v_0$. Thus,
    \[ \phi^* u = F(v) = F(\phi^* v_0) = \phi^* F(v_0)\]
    and hence $u = F(v_0)$ by tautness of $\Y$.
\end{proof}


Recall now that a monomorphism of categories is a functor which is injective on objects and faithful; similarly, a monomorphism of vf-categories is a vf-functor which is injective on points and vf-faithful. Therefore, we could define a \emph{sub-vf-category} of a vf-category $\X$ as a vf-category $\Y$ together with a vf-faithful vf-functor $\Y \to \X$ which is also injective on objects. However, just as how in ordinary category theory we are usually more interested in \emph{full} subcategories, here we also care about a more restrictive notion of sub-vf-category, which we call \emph{subspace}. For instance, tautness will be preserved under subspaces, but not under arbitrary sub-vf-categories.

\begin{defn}
A \emph{subspace} of $\X$ is a vf-category $\Y$ endowed with a vf-fully-faithful functor $\Y \mono \X$ which is also injective on points.
\end{defn}

Concretely, a subspace $\Y$ of $\X$ is determined by a subclass $\Y_0 \subseteq \X_0$ of points, together with the vf-arrows, identities, composites, and reindexings inherited from $\X$. Notationally, we then also write $\Y \subseteq \X$.

\begin{lem}
    Subspaces of taut vf-categories are taut.
    \begin{proof}
        Let $\Y \subseteq \X$ be a subspace of a taut vf-category $\X$. To show that $\Y$ is taut, let $u \colon p \vuto\tau q_{\phi t}$ be a $\phi$-reducible vf-arrow in $\Y$. Since $\Y$ is a subspace of $\X$, $u$ is also a vf-arrow in $\X$, and it is also $\phi$-reducible in $\X$. By tautness of $\X$, there exists a unique vf-arrow $u_0 \colon p \vuto\sigma q_s$ in $\X$ such that $u = \phi^* u_0$. Since both $p$ and $(q_s)_{s\in S}$ lie in $\Y$, it follows that $u_0$ is also a vf-arrow in $\Y$, and hence the claim.
    \end{proof}
\end{lem}


\begin{defn}\label{defn:open-subspace}
A subspace $\Y \subseteq \X$ is \emph{open} if for any point $p$ in $\Y$, any $S$-family of points $(q_s)_{s\in S}$ in $\X$, and any filter $\sigma$ on $S$, the class $\Hom^{\X}_\sigma (p, q_s)$ being inhabited implies that $q_s \in \Y$ for $s:\sigma$. A \emph{neighborhood} of a point $p$ in $\X$ is a subspace of $\X$ containing an open subspace $\Y\subseteq \X$ such that $p \in \Y_0$. 
\end{defn}

We denote by $\Open (\X)$ the (possibly large) poset of open subspaces of $\X$. If $\X$ is small, then $\Open(X)$ is a frame, and it determines a topology on $\X_0$: with respect to these topologies, the underlying maps of points of vf-functors are continuous. This construction gives thus rise to a 2-functor $\vfCat \to \TopSp$ which, with a slight abuse of notation, we also denote by $\Open$.


\begin{rmk}
Denote by $\bf{\Omega}$ the \emph{vf-category of truth values}, i.e.\ the vf-poset structure defined on the frame $\Omega \coloneqq \mathcal P(1)$ by setting $\alpha \vuleq{\sigma} \beta_s$ if and only if $\alpha \subseteq \int_{s:\sigma} \beta_s$, where 
\[ \int_{s:\sigma} \beta_s \coloneqq \bigcup_{S_0 \in \sigma^\op} \bigcap_{s\in S_0} \beta_s.\]
Then, a subspace $\Y\subseteq \X$ is open if the characteristic map $\X_0 \to {\Omega}$ of $\Y_0 \subseteq \X_0$ can be endowed with the structure of a vf-functor $\X \to \mathbf{\Omega}$.

Assuming classical logic, $\bf{\Omega}$ can be identified with the \emph{Sierpinski vf-category} $\mathbf{2}$. \todo{write this better}
\end{rmk}




\subsection{Tautness for vf-posets}As remarked in \Cref{ex:top-sp-are-taut-vf-posets}, topological spaces are taut as vf-posets. In this subsection, we show that this association yields an equivalence of 2-categories between topological spaces and (small) taut vf-posets, where the inverse construction is determined by the notion of openness of \Cref{defn:open-subspace}. Let us start by showing how to canonically associate a taut vf-poset to each vf-category.

\begin{cons}[The topological reflection]\label{cons:top-refl}
The \emph{topological reflection} $\Tref{\X}$ of a vf-category $\X$ is the vf-poset defined as follows:
\begin{enumerate}
    \item points of $\Tref{\X}$ are the same as those of $\X$;
    \item\looseness=-1 for each point $p$, $S$-family of points $(q_s)_{s\in S}$, and filter $\sigma$ on $S$, we set $p \vuleq{\sigma} q_s$ if and only if there exists a vf-arrow $p \vuto\tau q_{\phi t}$ in $\X$ for some cover $\phi \colon \tau \epi \sigma$.
\end{enumerate}
By construction, $\Tref{\X}$ is taut.
% In case $\X$ is itself a vf-poset, we refer to $\eta_{\X}$ as the \emph{tautification map}.

This construction realizes a 2-functor $\Tref{-} \colon \vfCAT\to \vfPOS$ which is left adjoint to the inclusion of vf-posets into vf-categories, and similarly for small vf-posets and vf-categories:
\[\begin{tikzcd}
	\vfPOS & \vfCAT
	\arrow[""{name=0, anchor=center, inner sep=0}, curve={height=12pt}, hook, from=1-1, to=1-2]
	\arrow[""{name=1, anchor=center, inner sep=0}, "{\Tref{-}}"', curve={height=12pt}, from=1-2, to=1-1]
	\arrow["\dashv"{anchor=center, rotate=-90}, draw=none, from=1, to=0]
\end{tikzcd}\qquad \begin{tikzcd}
	\vfPos & \vfCat
	\arrow[""{name=0, anchor=center, inner sep=0}, curve={height=12pt}, hook, from=1-1, to=1-2]
	\arrow[""{name=1, anchor=center, inner sep=0}, "{\Tref{-}}"', curve={height=12pt}, from=1-2, to=1-1]
	\arrow["\dashv"{anchor=center, rotate=-90}, draw=none, from=1, to=0]
\end{tikzcd}\]
In particular, since $\Tref{-}$ lands in taut vf-posets, the above reflections restrict to reflections of (small) taut vf-posets into (small) vf-categories.
\end{cons}


Both vf-category structures on $\X_0$, namely that of $\X$ and that of $\Tref{\X}$, detect a notion of open subspace: as we now show, in fact, the two coincide.

\begin{lem}
    For a vf-category $\X$, a subclass $\Y_0 \subseteq \X_0$ detects an open subspace of $\X$ if and only if it detects an open subspace of $\Tref{\X}$.
\begin{proof}
    Suppose that $\Y_0$ detects an open subspace $\Y$ of $\X$. To see that $\Y$ is also open as a subspace of $\Tref{\X}$, suppose that $p \vuleq{\sigma} q_s$ in $\Tref{\X}$ with $p \in \Y_0$. By definition, this means that there exists a vf-arrow $p \vuto{\tau} q_{\phi t}$ in $\X$ for some cover $\phi \colon \tau \epi \sigma$; by openness of $\Y$, this entails that $q_{\phi t} \in \Y_0$ for $t :\tau$. More explicitly:
    \[ \set{ t | q_{\phi t} \in \Y_0 } \in \tau \]
    from which, since $\phi$ is a cover in $\FF$:
    \[ \phi \left( \set{ t | q_{\phi t} \in \Y_0 }\right) = \set{ \phi t | q_{\phi t} \in \Y_0 } \in \sigma \]
    and hence $\set{ s | q_s \in \Y_0} \in \sigma$, meaning that $q_s \in \mathscr{Y}_0$ for $s: \sigma$. The converse implication is immediate.
\end{proof}
\end{lem}

In other words, the 2-functor $\Open \colon \vfCat \to \TopSp$ factors through taut vf-posets via the topological reflection $\Tref{-}$:
\[\begin{tikzcd}
	\TopSp && \vfPos_{\mathrm{taut}} && \vfCat
	\arrow[""{name=0, anchor=center, inner sep=0}, "\Tinj"{description}, curve={height=12pt}, hook, from=1-1, to=1-3]
	\arrow[curve={height=34pt}, hook, from=1-1, to=1-5]
	\arrow[""{name=1, anchor=center, inner sep=0}, "{{\Open}}"{description}, curve={height=12pt}, from=1-3, to=1-1]
	\arrow[""{name=2, anchor=center, inner sep=0}, curve={height=12pt}, hook, from=1-3, to=1-5]
	\arrow["\Open"', curve={height=34pt}, from=1-5, to=1-1]
	\arrow[""{name=3, anchor=center, inner sep=0}, "{\Tref{-}}"{description}, curve={height=12pt}, from=1-5, to=1-3]
	\arrow["\dashv"{anchor=center, rotate=-90}, draw=none, from=1, to=0]
	\arrow["\dashv"{anchor=center, rotate=-90}, draw=none, from=3, to=2]
\end{tikzcd}\]
As we now show, the 2-adjunction on the left is in fact an equivalence, thus justifying the name ``topological reflection'' for $\Tref{-}$. Concretely, it suffices to show that $\Tinj (\Open Y) \cong Y$ for every taut vf-poset $Y$. The proof of this fact, \Cref{prop:taut-vf-posets-come-from-spaces}, will make use of the following result: intuitively, it expresses that taut vf-posets admit `canonical' vf-arrows, corresponding to \emph{neighborhood filters}, which are \emph{minimal} with respect to reindexing.

\begin{prop}\label{prop:canonical-convergence-neighborhoods}
    Let $Y$ be a taut vf-poset and let $p\in Y_0$. Then, there exists a filter $\nu_p$ on $Y_0$ such that:
    \begin{enumerate}
        \item $p \vuleq{\nu_p} y$, where the family $y \colon \nu_p \hookrightarrow Y_0$ is represented by the identity function;
        \item $p \vuleq{\sigma} q_s$ holds in $Y$ if and only if $q \colon \sigma \to Y_0$ defines an $\FF$-arrow $\sigma \to \nu_p$.%  satisfying $(q_s)_{s:\sigma} = \phi^*(y)_{y:\nu_p}$.
    \end{enumerate}  
    Moreover, $\nu_p$ can be described as the filter of {neighborhoods} of $p$ in $Y$.
\begin{proof}
    First, we construct a filter $\nu_p$ with the desired properties. Let $N$ be the set of filters $\nu$ on $Y_0$ such that $p \vuleq{y:\nu} y$: note that $N$ is inhabited at least by the principal filter $\delta_p$ at $p\in Y_0$, as witnessed by the diagonal vf-arrow $\diag^{\delta_p}_p$.  Every filter $\nu \in N$ comes equipped with a canonical $\FF$-arrow $\nu \hookrightarrow Y_0$, represented by the identity function; thus, we also get a canonical $\FF$-arrow $\coprod_{\nu \in N} \nu\to Y_0$. Let $\nu_p$ be the filter on $Y_0$ arising via its image factorization.
    \[\begin{tikzcd}[sep = small]
    	{\coprod_{\nu \in N} \nu } && {Y_0} \\
    	& {\nu_p}
    	\arrow[from=1-1, to=1-3]
    	\arrow[two heads, from=1-1, to=2-2]
    	\arrow["y"', hook, from=2-2, to=1-3]
    \end{tikzcd}\]
    That $p \vuleq{y:\nu_p} y$ follows then by tautness, since $p \vuleq{y:\nu} y$ for each $\nu \in N$ implies that $p \vuleq{(\nu,y) : \coprod_\nu \nu} y$ by merging, and hence $p \vuleq{y:\nu_p} y$ as reducibility trivializes in vf-posets.

    To show property (ii), suppose that $p \vuleq{\sigma} q_s$ for some $S$-family $(q_s)_{s\in S}$ and let $q_*(\sigma)$ be the filter in the image factorization of $q \colon \sigma \to Y_0$. By tautness, $p \vuleq{\sigma} q_s$ entails that $p \vuleq{q_*(\sigma)} y$, i.e.\ that $q_*(\sigma) \in N$. Therefore, considering the inclusion $q_*(\sigma) \mono \coprod_{\nu\in N} \nu$, into the coproduct, we have that the following diagram commutes:
\[\begin{tikzcd}
	\sigma &&& \\
	{q_*(\sigma)} & {\coprod_{\nu \in N}\nu } && {Y_0} \\
	&& {\nu_p}
	\arrow[two heads, from=1-1, to=2-1]
	\arrow["q", curve={height=-18pt}, from=1-1, to=2-4]
	\arrow[from=2-1, to=2-2, hook]
	\arrow[bend left=20, hook, from=2-1, to=2-4]
	\arrow[curve={height=12pt}, from=2-1, to=3-3]
	\arrow[from=2-2, to=2-4]
	\arrow[two heads, from=2-2, to=3-3]
	\arrow["y"', hook, from=3-3, to=2-4]
\end{tikzcd}\]
In particular, the composite $\FF$-arrow $\sigma \to \nu_p$ is represented by $q$. % above satisfies $(q_s)_{s:\sigma} = \phi^*(y)_{y:\nu_p}$.

Finally, we describe $\nu_p$ explicitly as the filter of neighborhoods of $p$.
    \begin{itemize}
        \item Let $U \subseteq Y$ be an open neighborhood of $p$. Then, since $p \vuleq{\nu_p} y$ by property (i), by openness it follows that $y \in U_0$ for $y:\nu_p$, i.e.\ $U_0 \in \nu_p$.
        \item Conversely, let $V \subseteq Y$ be a subspace such that $V_0 \in \nu_p$, which equivalently means that $y \in V_0$ for $y: \nu_p$. Note that, by property (ii), $p \vuleq{\delta_p} p$ entails that the identity function realizes an $\FF$-arrow $\delta_p \to \nu_p$; thus, it induces a function $V^{\nu_p} \to V_0^{\delta_p}$. Since $(y)_{y:\nu_p} \in V_0^{\nu_p}$, it follows that $(y)_{y:\delta_p} \in V_0^{\delta_p}$, which is equivalent to $p \in V_0$.
        
        Consider then the subspace $U \subseteq V$ defined by:
        \[ U_0 \coloneqq \set{ q \in V_0 | V_0 \in \nu_q } \]
        By the above, $p \in U_0$; we now show that $U$ is open in $Y$, which proves that $V$ is a neighborhood of $p$. Suppose that $q \vuleq{\sigma} r_s$ in $Y$ with $q \in U_0$: we need to show that $r_s \in U_0$ for $s:\sigma$, which explicitly means that $\set{ s | r_s \in U_0}\in \sigma$. Taking composites, we have 
        \[ q \vuleq{\sigma} (r_s \vuleq{\nu_{r_s}} y ), \mbox{ that is, }q \vuleq{\sum_{s:\sigma}\nu_{r_s}} y,\]
        where now $y$ also denotes the projection $\sum_{s:\sigma}\nu_{r_s} \to Y_0$.  By property (ii) we then know that $y$ also defines an $\FF$-arrow $\sum_{s:\sigma}\nu_{r_s} \to \nu_q$; since $q \in U_0$, which by construction means that $V_0\in \nu_q$ i.e.\ $(y)_{y:\nu_q} \in V^{\nu_q}_0$, applying the induced function $V_0^{\nu_q} \to V_0^{\sum_{s:\sigma} \nu_{r_s}}$ we conclude that 
        \[\set{ s | (y)_{y:\nu_{r_s}} \in V_0^{\nu_{r_s}} }  = \set{ s | V_0 \in \nu_{r_s}} = \set{ s | r_s \in U_0}\in \sigma.\]
    \end{itemize}  
\end{proof}
\end{prop}


\begin{prop}\label{prop:taut-vf-posets-come-from-spaces}
    Let $Y$ be a taut vf-poset. For every point $p$, $S$-family of points $(q_s)_{s\in S}$, and filter $\sigma$ on $S$, the following are equivalent:
    \begin{enumerate}
        \item $p \vuleq{\sigma} q_s$ in $Y$;
        \item $p \vuleq{\sigma} q_s$ in $\Tinj( \Open  Y)$, i.e.\ for every open neighborhood $U \subseteq Y$ of $p$ it holds that $q_s \in U_0$ for $s:\sigma$.
    \end{enumerate}
\begin{proof}
    \looseness=-1
    Clearly $(i)\Rightarrow(ii)$ by definition of openness, so suppose $(ii)$. By \Cref{prop:canonical-convergence-neighborhoods} we know that $q$ realizes an $\FF$-arrow $\sigma \to \nu_p$ where $\nu_p$ is the filter of neighborhoods of $p$ in $\Tinj(\Open Y)$. Let $\nu_p'$ be the filter of neighborhoods of $p$ in $Y$: if we show that the identity map realizes an $\FF$-arrow $\nu_p \to \nu_p'$, i.e.\ that $\nu_p' \subseteq \nu_p$, then by \Cref{prop:canonical-convergence-neighborhoods} we can conclude that $p \vuleq{\sigma} q_s$ in $Y$. In turn, $\nu_p' \subseteq \nu_p$ follows since  open subspaces of $Y$ are open in $\Open Y$ by definition, and the opens of any topological space $X$ are open as subspaces of $\Tinj (X)$.
\end{proof}
\end{prop}

\begin{cor}
    $\TopSp \simeq \vfPostaut$.
\end{cor}

\begin{rmk}
    From now on, we will typically identify topological spaces with their associated (taut) vf-posets, omitting the functor $\Tinj$ from the notation.
\end{rmk}

We conclude this subsection with a useful characterization of convergence in topological reflections of small vf-categories, which follows by the previous results simply observing that $\Tref{\X}$ is taut, and hence a topological space, for each small vf-category $\X$.

\begin{cor}\label{prop:convergence-in-topological-reflection-is-witnessed}
    Let $\X$ be a small vf-category. For every point $p$, $S$-family of points $(q_s)_{s\in S}$, and filter $\sigma$ on $S$, the following are equivalent:
    \begin{enumerate}
        \item $p \vuleq{\sigma} q_s$ in $\Tref{\X}$;
        \item for every open neighborhood $\Y \subseteq \X$ of $p$ it holds that $q_s \in \Y_0$ for $s:\sigma$.
    \end{enumerate}
\end{cor}


% \begin{cor}
% Let $\X$ be a small vf-category and let $p \in \X_0$. Let $\nu_p$ be the filter on $\X_0$ defined by the open neighborhoods of $p$. Then, $p \vuleq{x:\nu_p}x$ in $\Tref{\X}$. 
% \begin{proof}
%     By \Cref{prop:convergence-in-topological-reflection-is-witnessed}, it suffices to see that every open neighborhood $\Y$ of $p$ in $\X$ satisfies $(x)_{x:\nu_p} \in \Y_0^{\nu_p}$, which is trivial since $\Y_0$ lies in $\nu_p$.
% \end{proof}
% \end{cor}

\subsection{Points of the localic reflection}Throughout this subsection, fix a topos $\E$. Recall that its \emph{localic reflection} $\Lref{\E}$ is defined as the frame $\Sub_{\E}(1)$ of subterminal objects. There are then two vf-posets that we can canonically associate to $\E$:
\begin{enumerate}
    \item[i.]on one hand, as we saw in \Cref{ex:points-of-toposes}, we can consider its canonical vf-category of points $\pt(\E)$, and take its topological reflection $\Tref{\pt(\E)}$; 
    \item[ii.]\looseness=-1 on the other hand, we can consider its localic reflection $\Lref{\E}$, take its space of points $\pt(\Lref{\E})$, and endow the latter with the canonical vf-poset structure induced by its topology.
\end{enumerate}
Note then that there is an evident vf-functor $\Theta \colon \Tref{\pt (\E)} \to \pt(\Lref{\E})$, defined by mapping each point $p \colon \E \to \Set$ to its restriction $p|_{\Sub_{\E}(1)} \colon \Sub_{\E}(1) \to \Omega$ to subterminal objects, which is a frame homomorphism and hence a point of $\Lref{\E}$. The goal of this subsection is to show that this vf-functor is in fact an \emph{equivalence} of vf-posets.

\red{we're working under the assumption that it is bijective on objects, so that vf-ff implies equivalence: but why is it so?}

First, we show that $\Theta$ is vf-fully-faithful. As it is trivially vf-faithful and its codomain $\pt(\Lref{\E})$ is taut, by \Cref{lem:vf-fully-faithful-equiv} it suffices to show that $\Theta$ is vf-full. To this end, we will make use of a choice-like result (\Cref{prop:section-up-to-filter}) which we can prove constructively by taking appropriate reduced powers; the core of the argument is enucleated by the following lemma.


\begin{lem}\label{lem:choice-up-to-filter}
    Let $(A_i)_{i\in I}$ be a family of inhabited sets. Then, there exists a proper filter $\lambda$ such that $\prod_{i\in I} A_i^\lambda$ is inhabited.
\begin{proof}
    Denote by $\mathcal P_\omega I$ the set of finite subsets of $I$ and let $L$ be the set $\sum_{I_0 \in \mathcal P_\omega I} \prod_{i\in I_0} A_i$ of all partial choice functions $(l_i)_{i\in I_0}$ for the family $(A_i)_{i\in I}$ defined on a finite domain $I_0$. For each $i \in I$, consider the set 
    \[ L_i \coloneqq \set{ l \in L | i \in \dom (l)}\]
    of partial functions defined at $i$. Note that any finite intersection $L_{i_0} \cap \dots \cap L_{i_n}$ is inhabited since so is each $A_i$; therefore, the filter $\lambda$ generated by the sets $(L_i)_{i\in I}$ is proper. By construction, for each $i\in I$, the family $(l_i)_{l\in L_i}$ represents an element $(l_i)_{l:\lambda}$ of $A_i^\lambda$; an element of $\prod_{i\in I} A_i^\lambda$ is thus given by $((l_i)_{l: \lambda})_{i\in I}$.
\end{proof}
\end{lem}

\looseness=-1
Recall now that a small category $\mathcal C$ is \emph{cofiltered} if every finite diagram in $\mathcal C$ admits a cone.

\begin{prop}\label{prop:section-up-to-filter}
    Let $\mathcal C$ be a small cofiltered category and let $A \colon \mathcal C \to \Set$ be a functor such that $A(c)$ is inhabited for each $c$ in $\mathcal C$. Then, there exists a proper filter $\lambda$ such that the functor $A^\lambda \colon \mathcal C \to \Set$ defined by $A^\lambda(c) \coloneqq A(c)^\lambda$ admits a global section. 
\begin{proof}
    First, note that it suffices to prove the claim under the additional hypothesis of $\prod_{c\in \Ob\mathcal C} A(c)$ being inhabited. Indeed, this `weaker' version can be applied to some reduced power $A^\lambda \colon \mathcal C \to \Set$ with respect to a proper filter $\lambda$, given by applying \Cref{lem:choice-up-to-filter} to the family $(A(c))_{c\in \mathcal C}$, obtaining that $(A^\lambda)^{\lambda'} \cong A^{\lambda' \times \lambda}\colon \mathcal C \to \Set$ admits a global section for some proper filter $\lambda'$. 
    
    Fix then an element $(a_c)_{c\in \mathcal C}$ of $\prod_{c\in \mathcal C} A(c)$. Consider the set:
    \[ L \coloneqq \sum_{I_0 \in \mathcal P_\omega (\Ob\mathcal C)} \prod_{c \in I_0} \Ob(\mathcal C/c),\] 
    whose elements are tuples $(l_c \colon d_c \to c)_{c\in I_0}$ of arrows in $\mathcal C$ for some finite set $I_0$ of objects in $\mathcal C$. For each $f \colon c \to c'$ in $\mathcal C$, consider the set:
    \[ L_f \coloneqq \set{ l \in L | c,c' \in \dom(l),\ d_c = d_{c'} \mbox{ and } f\circ l_{c}=  l_{c'}}.\]
    Note that any finite intersection $L_{f_0} \cap \dots \cap L_{f_n}$ is inhabited since $\mathcal C$ is cofiltered; therefore, the filter $\lambda$ generated by the sets $(L_f)_{f\in \operatorname{Arr}(\mathcal C)}$ is proper. 
    
    For each $c$ in $\mathcal C$ we have $L_{\operatorname{id}_c} = \set{ l \in L | c \in \dom (l)}$, and by construction the family $(Al_c(a_{d_c}))_{l \in L_{\operatorname{id}_c}}$ represents an element $(Al_c(a_{d_c}))_{l:\lambda}$ of $A^\lambda(c)$. To see that the association $c \mapsto (Al_c(a_{d_c}))_{l:\lambda}$ yields a global section of $A^\lambda$, we need to show its naturality, so let $f \colon c \to c'$ be any arrow in $\mathcal C$. Then, for each $l \in L_{f} \subseteq L_{\operatorname{id}_c} \cap L_{\operatorname{id}_{c'}}$, by definition we have:
    \[ Af ( Al_c(a_{d_c}) ) = A(f\circ l_c) (d_c) = Al_{c'} (a_{d_c}) = Al_{c'} (a_{d_{c'}}).\]
\end{proof}
\end{prop}

{\color{darkred}say somewhere: 

For a point $p \colon \E \to \Set$ of $\E$, we denote by $E_p$ the set $p(E)$.

For an object $E$ in $\E$, we denote by $|E|$ the subterminal object obtained by the image factorization of $E \to 1$.

For each family $(q_s)_{s\in S}$ of points of $\E$ and each filter $\sigma$ on the set $S$, the functor $\E \to \Set$ mapping $E \mapsto \int_{s:\sigma} E_{q_s}$ preserves epimorphisms.

}

\begin{prop}\label{prop:theta-is-vf-full}
    The vf-functor $\Theta \colon \Tref{\pt(\E)} \to \pt(\Lref{\E})$ is vf-full.
\begin{proof}
    Fix $p$ and $(q_s)_{s\in S}$ points of $\E$ and, for a filter $\sigma$ on $S$, suppose that $p \vuleq{\sigma} q_s$ in $\pt(\Lref{\E})$. Concretely, we need to prove that there exist a cover $\phi \colon \tau \epi \sigma$ in $\FF$ and a vf-arrow $p \vuto{t:\tau} q_{\phi t}$ in $\pt(\E)$, i.e.\ a natural transformation $p \Rightarrow \int_{t:\tau} q_{\phi t}$.

    Fix a small dense subcategory $\mathcal G$ of $\E$ which, without loss of generality, we can assume to admit finite limits.\footnote{Constructively, we can suppose that $\E$ is given as the category of sheaves on a lex site, so that $\mathcal G$ can be taken as its underlying category.} By density and by cocontinuity of $p$ \UT{here add ref to kelly as i did in my paper because typically people want both functors to be cocontinuous}, to define a natural $p \Rightarrow \int_{t:\tau} q_{\phi t}$ it suffices to define one between the restrictions $p\vert_{\mathcal G} \Rightarrow \int_{t:\tau} q_{\phi t}\vert_{\mathcal G}$. 
    
    Let $\mathcal C$ be the category of elements of the restriction $p\vert_{\mathcal G} \colon \mathcal G \to \Set$, which is cofiltered as the latter preserves finite limits, and consider the functor $A \colon \mathcal C \to \Set$ defined by $A(E, \xi) \coloneqq \int_{s:\sigma} E_{q_s}$. Note that, for each pair $(E,\xi)$ in $\mathcal C$, since $\xi \in E_p$, the set  $\int_{s:\sigma} E_{q_s}$ is inhabited, as can be seen from the following diagram of surjections and inclusions:
    \[\begin{tikzcd}[sep =small]
        {E_p} & {\int_{s:\sigma}E_{q_s}} \\
        {|E|_p} & {\int_{s:\sigma}|E|_{q_s}}
        \arrow[two heads, from=1-1, to=2-1]
        \arrow[two heads, from=1-2, to=2-2]
        \arrow["\subseteq"{description}, draw=none, from=2-1, to=2-2]
    \end{tikzcd}\]
    Thus, by \Cref{prop:section-up-to-filter}, we know that there exists a proper filter $\lambda$ such that the functor $A^\lambda \colon \mathcal C \to \Set$ admits a section. Note that, for each $(E,\xi)$ in $\mathcal C$, the set $A^\lambda(E,\xi)$ can be identified with the reduced product $\int_{(l,s):\lambda\cdot \sigma} E_{q_s}$ (cf.\ \Cref{rmk:monoidal-structure-on-F}). In other words, setting $\tau \coloneqq \lambda \cdot \sigma$, the functor $A^\lambda$ can be identified with the functor $(E,\mathcal C) \mapsto \int_{t:\tau} E_{q_{\phi t}}$ where $\phi \colon \tau \epi \sigma$ is the canonical map from the dependent product, which is a cover since $\lambda$ is proper. The section of $A^\lambda$ determined by \Cref{prop:section-up-to-filter} thus amounts to a natural transformation $p\vert_{\mathcal G} \Rightarrow \int_{t:\tau} q_{\phi t}\vert_{\mathcal G}$, which concludes the proof.
\end{proof}
\end{prop}

\UT{i wonder if density is problematic constructively..}